Symmetric Circuits for Rank Gates

### 2018-02-11 sections/background.tex
feb 11

\subsection{First Order and Fixed-Point Logic}
Let $L$ be a logic and $\tau$ be a vocabulary, we let $L[\tau]$ denote the set
of all $\tau$-formulas of $L$ (i.e. the set of all $L$-formulas with respect to
$\tau$). Let $\FO[\tau]$ denote \emph{first-order logic} with respect to the
vocabulary $\tau$. A formula in $\FO[\tau]$ is formed from atomic formulas, each
formed using variables from some countable sequence of (first-order) variable
symbols $(x, y, \ldots)$, the relation symbols in $\tau$, and the equality
symbol $=$, and then closing the set of atomic formulas under the Boolean
connectives, negation, and universal and existential quantification (i.e.
$\land$, $\lor$, $\neg$, $\forall$, and $\exists$).

@@

Let $L$ be a logic and $\tau$ a vocabulary. For $\phi \in L[\tau]$ and a
sequence of variables $\vec{x}$, we write $\phi(vec{x})$ if the free variables
in $\phi$ are among those in $\vec{x}$. Suppose $\vec{x} = (x_1, \ldots , x_k)$,
and let $\mathcal{A}$ be a $\tau$-structure with universe $U$, and let $\vec{a}
\in U^k$. We write $\mathcal{A} \models_L \phi[\vec{a}]$ if, and only if, $\phi$
is true in $\mathcal{A}$ with respect to $L$ where for all $i \in [k]$ each
variable $x_i$ is assigned to $a_i$. If the logic is obvious from context we
omit the subscript and just write $\mathcal{A} \models \phi[\vec{a}]$.

\subsection{Fixed-Point with Counting}
Let $\FPC[\tau]$ denote \emph{fixed-point logic with counting} with respect to
the vocabulary $\tau$. The logic $\FPC$ is an extension of $\FP$ that includes a
counting operator that allows the logic to determine the size of a definable
set, as well as other mechanisms for reasoning about quantities. We now briefly
describe the structure of a formula of $\FPC[\tau]$. In order to facilitate
counting we define terms which are assigned to natural numbers. Every
first-order variable is assigned a type, and so each variable is either a
\emph{vertex} or \emph{number} variable. Vertex variables are assigned to
elements of the universe of the structure, and play a similar role to the
variables that appear in formulas in $\FO[\tau]$ or $\FP[\tau]$. Number
variables are assigned to elements of $\nats$. We usually use Latin letters
(e.g. $x, y, z, \ldots$) to denote vertex variables and Greek letters (e.g.
$\mu, \nu, \epsilon, \ldots$) to denote number variables. All atomic formulas in
$\FP[\tau]$ are atomic formulas in $\FPC[\tau]$. We say that $t$ is a
\emph{number term} if $t$ a number variable or if $t$ is an application of the
\emph{counting operator}, i.e. $t = \#x. \phi(x)$, where $\phi \in \FPC[\tau]$
and $x$ is a vertex variable free in $\phi$. If $t_1$ and $t_2$ are number terms
then $t_1 \leq t_2$ and $t_1 = t_2$ are atomic formulas. The formulas of $\FPC$
are then formed by closing the set of atomic formulas under the usual Boolean
connectives, the first-order quantifiers (for vertex variables), and the
inflationary fixed-point operator. The second-order variables, such as those
that appear in an application of the fixed-point operator, are allowed to have
mixed-type, i.e. we associate with a second-order variable $R$ a type $(k, l)
\in \nats^2$ intended to mean that $R$ is assigned to a subset of $U^k \times
\nats^l$.

In order to ensure that the data complexity of $\FPC$ is in polynomial time, we
only allow bounded quantification when using number variables. In other words,
if $Q \in \{\forall, \exists \}$, $\phi \in \FPC[\tau]$, $\mu$ is a number
variable free in $\phi$, and $t$ is a number term, then $Q \mu \leq t . \phi$ is
a formula of $\FPC[\tau]$. Similarly, if $\phi$ is a number term then $\# \mu
\leq t .\phi$ is a number term in $\FPC[\tau]$.

If $\mathcal{A}$ is a $\tau$-structure with universe $U$, the counting term
$\#x. \phi(x)$ evaluates to $\vert \{a \in U : \mathcal{A} \models \phi[a]\}
\vert$ in $\mathcal{A}$, assuming a fixed assignment to the other free variables
in $\phi$. The intended semantics for bounded quantification is that the
variable may only be assigned to a value at most equal to the the evaluation of
the number term bounding the variable. The semantics of the $=$ and $\leq$
symbols are defined for number terms in the obvious manner. For a more detailed
description of the syntax and semantics of $\FPC$ please
see~\cite{grohe2017descriptive} or the introduction of~\cite{GradelP15a}.

@@

We now briefly describe the structure of a formula in $\FPR[\tau]$. For more
detail on the syntax and semantics of $\FPR$ please see~\cite{GradelP15a}
and~\cite{Dawar09logicswith}. As in the case of $\FPC$, we suppose the
first-order variable are assigned one of two types. We call the first type
\emph{vertex variables} and the second \emph{number variables} variables. Vertex
variables are assigned to elements of the universe and number variables are
assigned to natural numbers. All atomic formulas in $\FP[\tau]$ are atomic
formulas in $\FPC[\tau]$. We say that $t$ is a \emph{number term} if $t$ is a
number variable or if $t$ is an application of the \emph{rank operator}, i.e. $t
= [\rank (\vec{x}, \vec {\nu} \leq \vec{t}, \vec{y}\vec{\mu} \leq \vec{s}, \pi
\leq \eta). \phi]$, where $\phi$ is a number term, $\vec{t}$ and $\vec{s}$ are
tuples of number terms bounding the sequences of number variables $\vec{\mu}$
and $\vec{\nu}$, and $\eta$ is a number term bounding the number variable $\pi$.
If $t_1$ and $t_2$ are number terms then $t_1 \leq t_2$ and $t_1 = t_2$ are
atomic formulas. The formulas of $\FPR$ are then formed by closing the set of
atomic formulas under the usual Boolean connectives, the first-order
quantifiers, and the inflationary fixed-point operator. As in the case of
$\FPC$, the second-order variables may have mixed-type and we only allow bounded
quantification for number variables. We note additionally that formulas can be
treated as number terms that evaluate to zero or one.

We now briefly explain the intended semantics of the rank operator. Let $\phi
(\vec{x}, \vec {\nu}, \vec{y}\vec{\mu})$ be a number term, where $\vec{x}$ and
$\vec{y}$ are tuples of vertex variables, $\vec{\nu}$ and $\vec{\mu}$ are tuples
of number variables. Let and $\vec{s}$ and $\vec{t}$ be tuples of number terms
that bound the ranges of $\vec{\nu}$ and $\vec{\mu}$, respectively. Let $\pi$ be
a number variable and $r$ be a number term bounding the range of $\pi$. Let
$\mathcal{A}$ be a structure with universe $A$, and suppose we have a given
assignment to the all of the free variables appearing in $r$ and in the terms in
$\vec{t}$ and $\vec{s}$. Let $\mathbb{N}^{\leq\vec{t}} := \{\vec{n} \in
\mathbb{N}^{\vert \vec{v} \vert}: n_i \leq t^{\mathcal{A}}_i\}$, and let
$\mathbb{N}^{\leq \vec{s}}$ be defined similarly. Let $I:= A^{\vert \vec{x}
  \vert} \times \mathbb{N}^{\leq \vec{t}}$ and $J := A^{\vert \vec{y} \vert}
\times \mathbb{N}^{\leq \vec{s}}$. Let $M^{\mathcal{A}}_\phi : I \times J
\rightarrow \mathbb{N}$ be defined by $M^{\mathcal{A}}_\phi (\vec{a}\vec{n},
\vec{b}\vec{m}):= \phi^{\mathcal{A}}(\vec{a}\vec{n},\vec{b}\vec{m})$. For the
structure $\mathcal{A}$ and assignment $\pi \mapsto p$, for $p \leq r^{A}$ prime,
$[\rank (\vec{x}, \vec {\nu} \leq \vec{t}, \vec{y}\vec{\mu} \leq \vec{s}, \pi
\leq r). \phi]$ evaluates to the rank of $M^{\mathcal{A}}_\phi$ mod $p$, where
the matrix is understood to have entries in $\mathbb{F}_p$.

@@

Let $\FPrk[\tau]$ denote \emph{fixed-point logic with rank quantifiers} with
respect to the vocabulary $\tau$. This logic, first introduced by Dawar et
al~\cite{Dawar09logicswith}, is an extension of fixed-point logic by \emph{rank
  quantifiers}. For a prime $p$ and $r$ a natural number, a rank quantifier
$\rank^r_p$ is defined such that $\rank^r_p\vec{x}\vec{y}. \phi$ is interpreted
as $[\rank (\vec{x}, \vec{y}, \pi) . \phi] \leq t$, where $\pi$ is assigned to
$p$ and $t$ evaluates to $r$. The arity of this quantifier is $\vert \vec{x}
\vert + \vert \vert{y} \vert$. We thus have a rank quantifier for every prime
$p$, natural number $r$ and arity $k$. Let $\mathcal{R}$ be the set of all such
quantifiers. Let $\FPrk[\tau]$ be the closure of $\FP[\tau]$ under the
quantifiers in $\mathcal{R}$.

@@

For more detail on the syntax and semantics of these logics with rank
quantifiers please see~\cite{Dawar09logicswith}. It is easy to see that $\FPrk
\leq \FPR$ and $\FOrk \leq FOR$. In fact, Dawar et al.~\cite{Dawar09logicswith}
shows that these inclusions are strict. Moreover, we note that $[\rank(\vec{x},
\vec{y}). (\vec{x} = \vec{y}, \pi \leq r) \land \phi(\vec{x})]$ evaluates, for a
structure $\mathcal{A}$ and any assignment $\pi \mapsto p$, to the number of
assignments to $\vec{x}$ for which $\phi$ holds (i.e.\ $\vert \{\vec{a} \in
U^{\vert \vec{x} \vert} : \mathcal{A} \models \phi[\vec{a}]\} \vert$). It
follows that $\FPC \leq \FPR$ and $\FOC \leq \FOR$. Dawar et
al.\cite{Dawar09logicswith} show that these inclusions are strict.

\subsection{Group Theory}
Let $X$ be a set, we write $\sym_X$ to denote the symmetric group on the set
$X$. For $n \in \nats$ we write $\sym_n$ to abbreviate $\sym_{[n]}$. We assume
the reader is familiar with the basic theory of group actions.

Let $G$ be a group and let $H \leq G$. Let $X$ be a set on which a group action
of $G$ is defined. This group action defines a natural action on
$\mathcal{P}(X)$ (and hence $\mathcal{P}(\ldots(\mathcal{P}(X)))$) defined by
$\sigma \cdot S := \{\sigma x : x \in S\}$ for $\sigma \in G$, $S \subseteq X$.
Let $S \subseteq X$. We call $\stab_G(S) = \{\sigma in G : \forall x \in S \,
\sigma x = x\}$ the \emph{stabiliser} of $S$ with respect to $G$. We call
$\setstab_G(S) = \stab_G(\{S\}) = \{\sigma \in G : \sigma S = S\}$ the
\emph{set-wise stabiliser} of $S$ with respect to $G$. We call $\orb_G(S) :=
\{\sigma S : \sigma \in G\}$ the \emph{orbit} of $S$ with respect to $G$. If $S$
is a singleton we omit the set bracers, e.g.\ we write $\orb_G(x)$ for
$\orb_G(\{x\})$. If $G = \sym_n$ for some $n \in \nats$ we replace the subscript
with the number $n$, e.g.\ we write $\orb_n(x)$ for $\orb_{\sym_n}(x)$. If the
group $G$ is obvious from context we omit the subscript, e.g.\ we write
$\orb(x)$ for $\orb_G(x)$.

\subsection{Boolean Functions, Bases, and Circuits}
A (finite) \emph{Boolean function} is a function of the form $f: \{0,1\}^X
\rightarrow \{0,1\}$, for some finite index set $X$. We often take $X = [n]$,
and think of $f$ as taking $n$-length strings as inputs. We call a Boolean
function $f: \{0,1\}^X \rightarrow \{0,1\}$ \emph{symmetric} if for all $\sigma
\in \sym_X$ and $\vec{a} \in \{0,1\}^X$, $f (\vec{a} \sigma) = f(\vec{a})$. It
follows that a Boolean function is symmetric if, and only if, the output of the
function is entirely determined by the number of ones in the input. We say a
function $f : \{0,1\}^{*} \rightarrow \{0,1\}$ is \emph{symmetric} if its
restriction to inputs of size $n$ is symmetric for all $n$.

@@

Let $\BB$ be a basis. A \emph{Boolean circuit} $C$ defined over $\BB$ is a
labelled directed acyclic graph (DAG) with designated input gates $x_1, \ldots,
x_n$, each with in-degree $0$, internal gates labelled by elements of $\BB$, and
a single internal gate with out-degree $0$ designated as the output gate. We say
a circuit with $n$ input gates has \emph{order $n$}. Let $C$ be a Boolean
circuit of order $n$ and let $\vec{a} := (a_1, \ldots, a_n) \in \{0,1\}^n$. The
evaluation of $C$ for the input $\vec{a}$ is denoted by $C[\vec{a}]$ and is
defined by induction on the structure of the circuit. We denote the evaluation
of a gate $g$ in $C$ for the input $\vec{a}$ by $C[\vec{a}](g)$. If $g$ is an
input gate, then $g = x_i$ and $C[\vec{a}](g) = a_i$. If $g$ is an internal gate
then $g$ is labelled by a symbol denoting a Boolean operation, and
$C[\vec{a}](g)$ is the result of applying that operation the string formed from
the evaluations of those gates input to $g$. If $g$ is the output gate, let
$C[\vec{a}] = C[\vec{a}](g)$.

The \emph{size} of a circuit $C$, denoted by $\vert C \vert$ is the number of
gates in the circuit. The \emph{depth} of a gate is the longest path from an
input gate to that gate, and the \emph{depth} if the circuit is the depth of the
output gate. The \emph{width} of a circuit is the maximum size of a set of gates
with the same depth.

@@

\subsection{Circuits for Structures}
Anderson and Dawar~\cite{AndersonD17} define a circuit that takes as input an
encoding of a finite relational structure. Let $\tau$ be a vocabulary, $\BB$ be
a basis of symmetric functions and $q, n \in \nats$. A \emph{circuit for
  structures} (or \emph{circuit}) $C$ is a DAG with a designated set of input
gates each labelled by a relation symbol $R \in \tau$ and a tuple $\vec{a} \in
[n]^{\arty(R)}$, internal gates each labelled by an element of $\BB$, and a
designated set of output gates, each labelled by a unique element of $[n]^q$.

Let $\mathcal{A}$ be a $\tau$-structure of size $n$ and fix a bijection $\gamma$
from the universe of $\mathcal{A}$ to $[n]$. This bijection defines an encoding
of $\mathcal{A}$ as a $\tau$-structure with universe $[n]$. This encoding
defines an assignment to the input gates of the circuit, assigning the gate
labelled by the relation symbol $R$ and tuple $\vec{a})$ to one if, and only if,
$\vec{a}$ is an element of the interpretation of $R$ in the encoding of
$\mathcal{A}$. The circuit may then be recursively evaluated, with the
evaluation of the output gates taken as the output of the circuit. We note that
an evaluation of the output gates (and hence the output of the circuit) is an
element of $\{0,1\}^{[n]^q}$. The circuit then associates with every
$\tau$-structure and bijection $\gamma$ a set of $q$-tuples in $\mathcal{A}$.

A circuit is called \emph{invariant} if its output does not depend on the choice
of $\gamma$. In this case the circuit defines a $q$-ary query of
$\tau$-structures. If $C$ is an invariant circuit with $q = 0$, and the circuit
has a single output gate and decides a property of $\tau$-structures. A circuit
is called \emph{symmetric} if every permutation on the universe $[n]$, each of
which induces a permutation on the input gates, extends to an automorphism of
the circuit. A symmetric circuit is necessarily invariant.

>>>>

\subsection{First Order and Fixed-Point Logic}
Let $\FO[\tau]$ denote \emph{first-order logic} with respect to the vocabulary
$\tau$. A formula in $\FO[\tau]$ is formed from atomic formulas, each formed
using variables from some countable sequence of (first-order) variable symbols
$(x, y, \ldots)$, the relation symbols in $\tau$, and the equality symbol $=$,
and then closing the set of atomic formulas under the Boolean connectives,
negation, and universal and existential quantification (i.e. $\land$, $\lor$,
$\neg$, $\forall$, and $\exists$).

@@

Let $L$ be a logic and $\tau$ a vocabulary. For $\phi \in L[\tau]$ and a
sequence of variables $\vec{x}$, we write $\phi(\vec{x})$ if the free variables
in $\phi$ are among those in $\vec{x}$. Suppose $\vec{x} = (x_1, \ldots , x_k)$,
and let $\mathcal{A}$ be a $\tau$-structure with universe $U$, and let $\vec{a}
\in U^k$. We write $\mathcal{A} \models_L \phi[\vec{a}]$ if, and only if, $\phi$
is true in $\mathcal{A}$ with respect to $L$ where for all $i \in [k]$ each
variable $x_i$ is assigned to $a_i$. If the logic is obvious from context we
omit the subscript and just write $\mathcal{A} \models \phi[\vec{a}]$.

\subsection{Fixed-Point with Counting}
Let $\FPC[\tau]$ denote \emph{fixed-point logic with counting} with respect to
the vocabulary $\tau$. The logic $\FPC$ is an extension of $\FP$ that includes a
counting operator that allows the logic to determine the size of a definable
set, as well as other mechanisms for reasoning about quantities. We now briefly
describe the structure of a formula of $\FPC[\tau]$. Every first-order variable
is assigned a type, and so each variable is either a \emph{vertex} or
\emph{number} variable. Vertex variables are assigned to elements of the
universe of the structure, and play a similar role to the variables that appear
in formulas in $\FO[\tau]$ or $\FP[\tau]$. Number variables are assigned to
elements of $\nats$. We usually use Latin letters (e.g. $x, y, z, \ldots$) to
denote vertex variables and Greek letters (e.g. $\mu, \nu, \epsilon, \ldots$) to
denote number variables. All atomic formulas in $\FP[\tau]$ are atomic formulas
in $\FPC[\tau]$. We say that $t$ is a \emph{number term} if $t$ a number
variable or if $t$ is an application of the \emph{counting operator}, i.e. $t =
\#x. \phi(x)$, where $\phi \in \FPC[\tau]$ and $x$ is a vertex variable free in
$\phi$. If $t_1$ and $t_2$ are number terms then $t_1 \leq t_2$ and $t_1 = t_2$
are atomic formulas. $\FPC[\tau]$ includes all formulas formed by closing the
set of atomic formulas under the usual Boolean connectives, the first-order
quantifiers (for vertex variables), and the inflationary fixed-point operator.
The second-order variables, such as those that appear in an application of the
fixed-point operator, are allowed to have mixed-type, i.e. we associate with a
second-order variable $R$ a type $(k, l) \in \nats^2$ intended to mean that $R$
is assigned to a subset of $U^k \times \nats^l$.

We also allow quantification over the number sort, but in order to ensure that
the data complexity of $\FPC$ is in polynomial time we only allow bounded
quantification. In other words, if $Q \in \{\forall, \exists \}$, $\phi \in
\FPC[\tau]$, $\mu$ is a number variable free in $\phi$, and $t$ is a number
term, then $Q \mu \leq t . \phi$ is a formula of $\FPC[\tau]$. Similarly, if
$\phi$ is a number term then $\# \mu \leq t .\phi$ is a number term in
$\FPC[\tau]$.

If $\mathcal{A}$ is a $\tau$-structure with universe $U$, the counting term
$\#x. \phi(x)$ evaluates to $\vert \{a \in U : \mathcal{A} \models \phi[a]\}
\vert$ in $\mathcal{A}$, assuming a fixed assignment to the other free variables
in $\phi$. The intended semantics for bounded quantification is that the
variable may only be assigned to a value at most equal to the the evaluation of
the number term bounding the variable. The semantics of the `$=$' and `$\leq$'
symbols are defined for number terms in the obvious manner. For a more detailed
description of the syntax and semantics of $\FPC$ please
see~\cite{grohe2017descriptive} or the introduction of~\cite{GradelP15a}.

@@

We now briefly describe the structure of a formula in $\FPR[\tau]$. For more
detail on the syntax and semantics of $\FPR$ please see~\cite{GradelP15a}
and~\cite{Dawar09logicswith}. As in the case of $\FPC$, each variable is either
a number or vertex variable, with vertex variables being assigned to elements of
the universe and number variables assigned to natural numbers. All atomic
formulas in $\FP[\tau]$ are atomic formulas in $\FPR[\tau]$. We say that $t$ is
a \emph{number term} if $t$ is a number variable or if $t$ is an application of
the \emph{rank operator}, i.e. $t = [\rank (\vec{x}, \vec {\nu} \leq \vec{t},
\vec{y}\vec{\mu} \leq \vec{s}, \pi \leq \eta). \phi]$, where $\phi$ is a number
term, $\vec{t}$ and $\vec{s}$ are tuples of number terms bounding the sequences
of number variables $\vec{\mu}$ and $\vec{\nu}$, and $\eta$ is a number term
bounding the number variable $\pi$. If $t_1$ and $t_2$ are number terms then
$t_1 \leq t_2$ and $t_1 = t_2$ are atomic formulas. The formulas of $\FPR$ are
then formed by closing the set of atomic formulas under the usual Boolean
connectives, the first-order quantifiers, and the fixed-point operator. As in
the case of $\FPC$, the second-order variables may have mixed-type and we only
allow bounded quantification for number variables. We note additionally that
formulas may be treated as number terms that evaluate to zero or one.

We now briefly explain the intended semantics of the rank operator. Let $\phi
(\vec{x}, \vec {\nu}, \vec{y}\vec{\mu})$ be a number term, where $\vec{x}$ and
$\vec{y}$ are tuples of vertex variables, $\vec{\nu}$ and $\vec{\mu}$ are tuples
of number variables. Let and $\vec{s}$ and $\vec{t}$ be tuples of number terms
that bound the ranges of $\vec{\nu}$ and $\vec{\mu}$, respectively. Let $\pi$ be
a number variable and $r$ be a number term bounding the range of $\pi$. Let
$\mathcal{A}$ be a structure with universe $A$, and suppose we have a given
assignment to the all of the free variables appearing in $r$ and in the terms in
$\vec{t}$ and $\vec{s}$. Let $\mathbb{N}^{\leq\vec{t}} := \{\vec{n} \in
\mathbb{N}^{\vert \vec{v} \vert}: n_i \leq t^{\mathcal{A}}_i\}$, and let
$\mathbb{N}^{\leq \vec{s}}$ be defined similarly. Let $I:= A^{\vert \vec{x}
  \vert} \times \mathbb{N}^{\leq \vec{t}}$ and $J := A^{\vert \vec{y} \vert}
\times \mathbb{N}^{\leq \vec{s}}$. Let $M^{\mathcal{A}}_\phi : I \times J
\rightarrow \mathbb{N}$ be defined by $M^{\mathcal{A}}_\phi (\vec{a}\vec{n},
\vec{b}\vec{m}):= \phi^{\mathcal{A}}(\vec{a}\vec{n},\vec{b}\vec{m})$. For the
structure $\mathcal{A}$ and assignment $\pi \mapsto p$, for $p \leq r^{A}$
prime, $[\rank (\vec{x}, \vec {\nu} \leq \vec{t}, \vec{y}\vec{\mu} \leq \vec{s},
\pi \leq r). \phi]$ evaluates to the rank of $M^{\mathcal{A}}_\phi$ mod $p$,
where the entries in $M^{\mathcal{A}}$ are interpreted as elements of
$\mathbb{F}_p$.

@@

Let $\FPrk[\tau]$ denote \emph{fixed-point logic with rank quantifiers} with
respect to the vocabulary $\tau$. This logic, first introduced by Dawar et
al~\cite{Dawar09logicswith}, is similar to $\FPR$, except that it extends $\FP$
by \emph{rank quantifiers}, rather than a rank operator, that can decide if a
definable matrix has a rank bounded by some constant. We define for each prime
$p$, and natural number $r$, a rank quantifier $\rank^r_p$, such that
$\rank^r_p\vec{x}\vec{y}. \phi$ is interpreted as $[\rank (\vec{x}, \vec{y},
\pi) . \phi] \leq t_r$, where $\pi$ is assigned to $p$ and $t_r$ is a number
term that evaluates to $r$. The arity of this quantifier is $\vert \vec{x} \vert
+ \vert \vec{y} \vert$. Let $\mathcal{R}$ be the set of all such quantifiers and
$\FPrk[\tau]$ be the closure of $\FP[\tau]$ under $\mathcal{R}$.

@@

For more detail on the syntax and semantics of logics rank quantifiers please
see~\cite{Dawar09logicswith}. It is easy to see that $\FPrk \leq \FPR$ and
$\FOrk \leq \FOR$. Moreover, we note that $[\rank(\vec{x}, \vec{y}).
(\vec{x} = \vec{y}, \pi \leq r) \land \phi(\vec{x})]$ evaluates, for a structure
$\mathcal{A}$ and any assignment $\pi \mapsto p$, to the number of assignments
to $\vec{x}$ for which $\phi$ holds (i.e.\ $\vert \{\vec{a} \in U^{\vert \vec{x}
  \vert} : \mathcal{A} \models \phi[\vec{a}]\} \vert$). It follows that $\FPC
\leq \FPR$ and $\FOC \leq \FOR$. Dawar et al.\cite{Dawar09logicswith} show that
all of these inclusions are strict.

\subsection{Group Theory}
Let $X$ be a set, we write $\sym_X$ to denote the symmetric group on the set
$X$. For $n \in \nats$ we write $\sym_n$ for $\sym_{[n]}$.

Let $G$ be a group and let $H \leq G$. Let $X$ be a set on which a group action
of $G$ is defined. This group action can be lifted to a group action on
$\mathcal{P}(X)$ (and hence $\mathcal{P}(\ldots(\mathcal{P}(X))\ldots)$) defined
by $\sigma \cdot S := \{\sigma x : x \in S\}$ for $\sigma \in G$, $S \subseteq
X$. Let $S \subseteq X$. We let $\stab_G(S) := \{\sigma \in G : \forall x \in S
\,, \sigma x = x\}$ be the \emph{stabiliser} of $S$ with respect to $G$. We let
$\setstab_G(S) := \stab_G(\{S\}) = \{\sigma \in G : \sigma S = S\}$ be the
\emph{set-wise stabiliser} of $S$ with respect to $G$. We let $\orb_G(S) :=
\{\sigma S : \sigma \in G\}$ be the \emph{orbit} of $S$ with respect to $G$. If
$S$ is a singleton we omit the set bracers, e.g.\ we write $\orb_G(x)$ for
$\orb_G(\{x\})$. If $G = \sym_n$ for some $n \in \nats$ we replace the subscript
with the number $n$, e.g.\ we write $\orb_n(x)$ for $\orb_{\sym_n}(x)$, and if
the group $G$ is obvious from context we omit the subscript entirely.

\subsection{Boolean Functions, Bases, and Circuits}
A \emph{Boolean function} is a function of the form $f: \{0,1\}^X \rightarrow
\{0,1\}$, for some finite index set $X$. We call a Boolean function $f:
\{0,1\}^X \rightarrow \{0,1\}$ \emph{symmetric} if for all $\sigma \in \sym_X$
and $\vec{a} \in \{0,1\}^X$, $f (\vec{a} \sigma) = f(\vec{a})$. It follows that
a Boolean function is symmetric if, and only if, the output of the function is
entirely determined by the number of ones in the input. We say a function $f :
\{0,1\}^{*} \rightarrow \{0,1\}$ is \emph{symmetric} if its restriction to
inputs of size $n$ is symmetric for all $n$.

@@

Let $\BB$ be a basis. A \emph{Boolean circuit} $C$ defined over $\BB$ is a
labelled directed acyclic graph (DAG) with designated input gates $x_1, \ldots,
x_n$, each with in-degree $0$, internal gates labelled by elements of $\BB$, and
a single internal gate with out-degree $0$ designated as the output gate. We say
a circuit with $n$ input gates has \emph{order $n$}. The evaluation of circuit
$C$ of order $n$ for the input $\vec{a} \in \{0,1\}^n$ is denoted by
$C[\vec{a}]$ and computed by recursively evaluation the gates in the circuit. We
denote the evaluation of a gate $g$ in $C$ for the input $\vec{a}$ by
$C[\vec{a}](g)$. If $g$ is an input gate, then $g = x_i$ for some $i \in [n]$,
and $C[\vec{a}](g) = a_i$. If $g$ is an internal gate then $g$ is labelled by a
symbol denoting a Boolean operation, and $C[\vec{a}](g)$ is the result of
applying that operation the string formed from the evaluations of those gates
input to $g$. If $g$ is the output gate, let $C[\vec{a}] = C[\vec{a}](g)$.

The \emph{size} of a circuit $C$, denoted by $\vert C \vert$, is the number of
gates in the circuit. The \emph{depth} of a gate $g$ is the longest path from an
input gate to $g$, and the \emph{depth} if the circuit is the depth of the
output gate. The \emph{width} of a circuit is the maximum size of a set of gates
with the same depth.

@@

\subsection{Circuits for Structures}
A \emph{circuit for structures} (or just \emph{circuit}), defined by Anderson
and Dawar~\cite{AndersonD17}, takes as input an encoding of a $\tau$-structure
and outputs a set of tuples from that structure. More formally, for a vocabulary
$\tau$, a basis $\BB$ of symmetric functions, and $q, n \in \nats$, a
\emph{circuit for structures} is a DAG with (i) a designated set of input gates,
each labelled by a relation symbol $R \in \tau$ and a tuple $\vec{a} \in
[n]^{\arty(R)}$, (ii) a set of internal gates, each labelled by an element of
$\BB$, and (iii) a designated set of output gates, each labelled by a unique
element of $[n]^q$.

Let $\mathcal{A}$ be a $\tau$-structure of size $n$ and $\gamma$ be a bijection
from the universe of $\mathcal{A}$ to $[n]$. The bijection $\gamma$ defines an
encoding of $\mathcal{A}$ as a $\tau$-structure with universe $[n]$. The input
gate labelled by the relation symbol $R$ and tuple $\vec{a}$ is assigned to one
if, and only if, $\vec{a}$ is an element of the interpretation of $R$ in the
encoding of $\mathcal{A}$. The circuit may then be recursively evaluated, and
the evaluation of the output gates is taken as the output of the circuit. Since
each output gate is labelled by an element of $[n]^q$, the output of the circuit
is an element of $\{0,1\}^{[n]^q}$. In other words, a circuit maps a
$\tau$-structure $\mathcal{A}$ and bijection $\gamma$ to a set of $q$-tuples in
$\mathcal{A}$.

A circuit is called \emph{invariant} if its output does not depend on the choice
of $\gamma$. In this case the circuit defines a $q$-ary query on
$\tau$-structures. If $C$ is an invariant circuit with $q = 0$, and the circuit
has a single output gate, then $C$ decides a property of $\tau$-structures. A circuit
is called \emph{symmetric} if every permutation on the universe $[n]$, each of
which induces a permutation on the input gates, extends to an automorphism of
the circuit. A symmetric circuit is necessarily invariant.
